---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cutnumberint

<!-- badges: start -->
[![R build status](https://github.com/yutannihilation/cutnumberint/workflows/R-CMD-check/badge.svg)](https://github.com/yutannihilation/cutnumberint/actions)
<!-- badges: end -->

A simple R package to provide only one function `cut_number_int()`, a nicer version of `ggplot2::cut_number()`.

## Installation

``` r
devtools::install_github("yutannihilation/cutnumberint")
```

## Example

`ggplot2::cut_number()` is one of the hidden gems in ggplot2. It's really neat, but I have two complaints about it. First, it cannot handle 0-inflated data.

```{r error_example, error=TRUE}
library(ggplot2)

set.seed(403)
x <- rpois(100, 0.5)

cut_number(x, n = 3)
```


Second one is a general complaint about `cut()`. For integer(-ish) data, I want intuitive labels like `6~10` instead of `(5.5,10]`.

```{r bad_labels, error=TRUE}
cut_number(1:10, 2)
```

`cut_number_int()` can handle these nicely.

```{r no_error}
library(cutnumberint)

cut_number_int(x, n = 3)
```

```{r no_error_plot}
d <- data.frame(x = cut_number_int(x, n = 3))
ggplot(d) +
  geom_bar(aes(x))
```

```{r nice_labels, error=TRUE}
table(cut_number_int(1:1000, 3))
```

### Caveats

For simplicity, this package handles only the cases where the inflated values are located on the lowest bound. If it appears on the middle of the data range, `cut_number_int()`.


```{r out_of_scope, error=TRUE}
cut_number_int(c(0, rep(1, 100), 2:9), 5)
```

If you need more powerful functions, probably you can find nicer binning packages on CRAN.

* <https://www.r-pkg.org/search.html?q=binning>
